Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-10 ms.tex
cleaned up2

Cooling instability can, however, be prevented if destabilizing radiative cooling ($\dot{q}_{\rm rad} \propto T^{-2.7}$) is overwhelmed by
other sources of cooling, namely the {\it stabilizing} term $\propto -q c_{s}^{2}
\propto -T$ in the entropy equation (eq.~[\ref{eq:dsdt}]).
Neglecting radiative cooling to first order, this term is balanced at the stagnation
radius by the external heating term, $\dot{q}_{\rm heat} = q
\tilde{v}_{w}^{2}/2$.  We assess thermal stability near the stagnation radius by comparing the ratio of external heating to
radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   630 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   660 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs})
for the gas and stellar densities at the stagnation radius,
respectively.  We have approximated the cooling function for $T <
2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6
  \text{K}\right)^{-0.7} $erg cm$^3 $s$^{-1}$, which assumes solar
metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).  Note that in
our runs with radiative cooling we take $\Lambda(T)$ to be a piecewise
function with $\Lambda = $. 
To within a constant of order unity, equation (\ref{eq:cooling2}) also
equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n
kT/2 \mu)/|\dot{q}_{\rm rad}|$ to the free-fall time $t_{\rm ff}$ at the
stagnation radius, as can be derived using the equality
\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations
\eqref{eq:mdot_analytic},\eqref{eq:mdot_gas}, and
\eqref{eq:rhostarrs}.  \citet{McCourt+12} argue cooling instability
develops in a hydrostatic atmosphere if $t_{\rm cool} \lesssim 10 t_{\rm ff}$,
so equation (\ref{eq:cooling2}) represents a good proxy for
instability in this case as well (see also \citealt{Li&Bryan14a}, \citealt{Li&Bryan14b}).

@@

Also note that for $v_{w}=600$ and $1200$ km s$^{-1}$ we have that
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \sim t_{\rm cool}/t_{\rm ff}$
near the stagnation radius.  However, the two ratios diverge from each
other at low heating ($v_{w} \lesssim 300$ km s$^{-1}$) because
equations (\ref{eq:mdot_gas}), (\ref{eq:rhors}) underestimate the true
gas density, which is greater in a subsonic inflow (weak heating) than
in a supersonic outflow (strong heating).

@@

Thermal conduction, by contrast, results in only minor changes to our solutions for our fiducial value of the saturation parameter $\phi = 0.1$.  Conduction may, however, offset the effects of radiative cooling in making the flow more thermally stable (e.g.~\citealt{Zakamska&Narayan03}).

@@

\item Heating from SMBH feedback, unlike stellar heating, depends on
  the SMBH accretion rate, which is itself a function of the heating
  rate.  Compton heating is generally unimportant compared to stellar
  wind heating for the average SFH, but it can dominate in the case of
  an impulsive star burst (Fig.~\ref{fig:vwSourcesImp}).  Although in
  principle Compton heating is sufficiently high to prevent thermal
  instabilities when acting alone, i.e. $\dot{M}_{C} \lesssim
  \dot{M}_{\rm TI}$ (Fig.~\ref{fig:bh_growth}), SMBH feedback may not
  however be capable of truly stabilizing the flow given its inability
  to respond instantaneously to local changes in the properties of the
  gas.  Kinetic feedback could also be important in determining
  $\dot{M}(M_{\bullet}$) when stellar heating is weak
  ($\S\ref{sec:kinetic}$) but is more challenging to quantify.

>>>>

Cooling instability can, however, be prevented if destabilizing radiative cooling ($\dot{q}_{\rm rad} \propto T^{-2.7}$) is overwhelmed by
other sources of cooling, namely the {\it stabilizing} term $\propto -q c_{s}^{2}
\propto -T$ in the entropy equation (eq.~[\ref{eq:dsdt}]).
Neglecting radiative cooling to first order, this term is balanced at the stagnation
radius by the external heating term, $\dot{q}_{\rm heat} = q
\tilde{v}_{w}^{2}/2$.  We assess thermal stability near the stagnation radius by comparing the ratio of external heating to
radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   630 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   660 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs})
for the gas and stellar densities at the stagnation radius,
respectively.  We have approximated the cooling function for $T <
2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6
  \text{K}\right)^{-0.7} $erg cm$^3 $s$^{-1}$, which assumes solar
metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).  Note that in
our runs with radiative cooling we take $\Lambda(T)$ to be a piecewise
function with $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6
  \text{K}\right)^{-0.7} $erg cm$^3 $s$^{-1}$ for $T\leq 2\times
10^{7}$ K and $\Lambda = 2.3 \times 10^{-24} \left(T/10^6
  \text{K}\right)^{0.5}$ for $T>2\times 10^{7}$ K.

To within a constant of order unity, equation (\ref{eq:cooling2}) also
equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n
kT/2 \mu)/|\dot{q}_{\rm rad}|$ to the free-fall time $t_{\rm ff}$ at the
stagnation radius, as can be derived using the equality
\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations
\eqref{eq:mdot_analytic},\eqref{eq:mdot_gas}, and
\eqref{eq:rhostarrs}.  \citet{McCourt+12} argue cooling instability
develops in a hydrostatic atmosphere if $t_{\rm cool} \lesssim 10 t_{\rm ff}$,
so equation (\ref{eq:cooling2}) represents a good proxy for
instability in this case as well (see also \citealt{Li&Bryan14a}, \citealt{Li&Bryan14b}).

@@

Also note that for $v_{w}=600$ and $1200$ km s$^{-1}$ we have that
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \sim t_{\rm cool}/t_{\rm ff}$
near the stagnation radius.  However, the two ratios diverge from each
other at low heating ($v_{w} \lesssim 300$ km s$^{-1}$) because
equations (\ref{eq:mdot_gas}), (\ref{eq:rhors}) underestimate the true
gas density, which is greater in a subsonic flow (weak heating) than
in a supersonic outflow (strong heating).

@@

Thermal conduction, by contrast, results in only minor changes to our solutions for our fiducial value of the saturation parameter $\phi = 0.1$.  Conduction can, however, offset the effects of radiative cooling in making the flow more thermally stable (e.g.~\citealt{Zakamska&Narayan03}).

@@

\item Unlike stellar heating, heating from SMBH feedback depends on
  the accretion rate, which itself depends on the heating.  Compton heating is generally unimportant compared to stellar
  wind heating for the average SFH, but it can dominate in the case of
  an impulsive star burst (Fig.~\ref{fig:vwSourcesImp}).  Although in
  principle Compton heating is sufficiently high to prevent thermal
  instabilities when acting alone, i.e. $\dot{M}_{C} \lesssim
  \dot{M}_{\rm TI}$ (Fig.~\ref{fig:bh_growth}), SMBH feedback may not
  however be capable of truly stabilizing the flow given its inability
  to respond instantaneously to local changes in the properties of the
  gas.  Kinetic feedback could also be important in determining
  $\dot{M}(M_{\bullet}$) when stellar heating is weak
  ($\S\ref{sec:kinetic}$) but is more challenging to quantify.
